\Large \bf Dynamic Program Dialysis with Lya

### 2020-01-06 paper/draft.tex
Update language features

\begin{table}[t]
\center
\footnotesize
\setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Language features used and their availability in other languages}.
    Broadly, the requirements of module dialysis can be broken down into to five elementary requirements.
  }
}
\begin{tabular*}{\columnwidth}{l @{\extracolsep{\fill}} llllllll}
\toprule
                                & Lua     & JavaScript  & Ruby   &      &      &      &      &         \\
\midrule
Instrument \ttiny{import}s      &         &             &        &      &      &      &      &         \\
Walk objects                    &         &             &        &      &      &      &      &         \\
Shadow variables                &         &             &        &      &      &      &      &         \\
Wrap values                     &         &             &        &      &      &      &      &         \\
Global names                    &         &             &        &      &      &      &      &         \\
\bottomrule
\end{tabular*}
\label{tab:compat}
\vspace{-5mm}
\end{table}

@@

First, \sys replaces the implementation of the import statement (\eg \ttt{require});
  this feature is important for compatibility reasons---otherwise developers would have to replace all import statements in a deeply nested tree.
If this is a function (\eg JavaScript, Scheme, Racket \etc) then replacement can be easily achieved by replacing its definition.
If this is a special-keyword statement such as \ttt{import}, this can be achieved by a rewriting rule that replaces import with a special function;
% This is simplified by the fact that the import syntax is trivial.

Second, \sys makes extensive use of weak maps to support to check for equality and, more frequently, for checking for cycles during object traversal (which requires equality).

Dynamic languages have features---\eg name (re-)binding, value introspection, dynamic code evaluation, and access interposition---that enable runtime transformations~\cite{aop, metaobject}.
They conveniently unify module identification with interposition: % used for performance monitoring and replica construction:
  a single function or function-like operator locates a module, interprets it, and applies transformations before exposing its interface in the caller context.

Other dynamic languages such as Lua, Ruby, and Python conveniently unify module identification with runtime interposition:
% used for performance monitoring and replica construction:
  a single function or function-like operator locates a module, interprets it, and applies transformations before exposing its interface in the caller context.
All these languages offer runtime capabilities such as name (re-)binding, value introspection, name interception, and access interposition.
% Languages such as Smalltalk and Lisp that treat ``everything'' as an object or a  function, need to interpose through object wrappers or function closures, respectively.

Compiled languages such as Haskell, OCaml, and Rust would apply transformations using compile-time metaprogramming facilities.
Examples include templates (\eg Template Haskell) and macro expansions (\eg Rust's macro system).
Such facilities would alleviate the runtime overhead of transformations, improving program startup times;
  however, they would not affect the overhead of contract evaluation, as this is still a runtime operation.
Static type checking could further aid developers by issuing warnings of incompatible privilege settings.
Contrary to OSMLC, there is no need for support from the runtime linker~\cite[\S VII]{breakapp:ndss:2018} because compartments do not execute as separate processes.

>>>>

\begin{table}[t]
\center
\footnotesize
\setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Language features used and their availability in other languages}.
    Broadly, the requirements of module dialysis can be broken down into to five elementary requirements.
  }
}
\begin{tabular*}{\columnwidth}{l @{\extracolsep{\fill}} llllllll}
\toprule
                                & Lua     & JavaScript  & Ruby   &      &      &      &      &         \\
\midrule
Instrument \ttiny{import}s      &         &             &        &      &      &      &      &         \\
Traverse objects                &         &             &        &      &      &      &      &         \\
Wrap values                     &         &             &        &      &      &      &      &         \\
(Weak) Maps                     &         &             &        &      &      &      &      &         \\
Shadow variables                &         &             &        &      &      &      &      &         \\
% Global names                    &         &             &        &      &      &      &      &         \\
\bottomrule
\end{tabular*}
\label{tab:compat}
\vspace{-5mm}
\end{table}

@@

First, \sys replaces the implementation of the import statement (\eg \ttt{require}).
% This feature is important for compatibility reasons---otherwise developers would have to replace all import statements in their applications.
If this statement is a function (\eg JavaScript, Scheme, Racket \etc) then replacement can be easily achieved by replacing its definition.
If this statement is a special non-function keyword such as \ttt{import}, this can be achieved by a rewriting rule that replaces import with a special function;
% This is simplified by the fact that the import syntax is trivial.

\sys also needs the ability to interpose on values and transform their components.
This is relatively easy in languages that provide meta-programming capabilities.
Dynamic programming languages offer runtime reflection and conveniently expose object accesses as (over-loadable) functions---\eg \ttt{Proxy} objects in JavaScript, metatables in Lua, metaclasses in Python, and direct-accessor metaprogramming Ruby.
Compiled languages such as Haskell, OCaml, and Rust would apply transformations with static metaprogramming facilities, traversing and transforming objects at compile time.
Examples include templates and macro expansions such as Haskell's Template Haskell and Rust's macro system.
And some languages, such as Haskell, Java, and Julia, provide both static and dynamic reflection capabilities, allowing considerable freedom in the choice of implementation.
% Fifth, \sys makes extensive use of interposition which is also available in different programming languages.
% Static programming languages provide different types of support for

\sys makes extensive use of weak maps to support to check for object equality.
Examples of uses include checking for cycles during object traversal and checking whether an object has already been wrapped.
Hash tables that are indexed by reference are supported in virtually all programming languages.
An additional feature useful in garbage-collected languages would be for these (copies of) references to not affect the reclamation of original objects.
Languages either support the ability to create such a weak reference from a normal reference (\eg Python's weak references), or provide structures that automatically hold weak versions of these references (\eg JavaScript's Weak Map).

Finally, \sys makes extensive use of variable shadowing to hide access to the original, unmodified versions of values.
This is a feature that is universally available in programming languages;\footnote{
  A notable exception is the scripting language CoffeeScript, which can mimic the same feature using function scope.
}
Deferences in scoping rules---\ie what the shadowed variable ends up referring to upon execution---do not affect the usefulness of this feature, although we have found lexical scoping easier than dynamic scoping.

To conclude, all major languages provide enough facilities to support the dialysis---albeit with a different level of convenience.
Dynamic languages have features---\eg name (re-)binding, value introspection, dynamic code evaluation, and access interposition---that enable runtime transformations.
Runtime transformations conveniently unify module identification with interposition:
  a single function or function-like operator locates a module, interprets it, and applies transformations before exposing its interface in the caller context.
Compiled languages such as Haskell, OCaml, and Rust apply transformations using compile-time metaprogramming facilities.
Compile time transformations alleviate much of the runtime overhead of transformations, improving program startup times;
  and type checking could further aid developers by issuing warnings of incompatible privilege settings.
